---
title: "PA1_template"
author: "trumayanvoid"
date: "5/17/2020"
output:
  html_document: 
    fig_caption: yes
    highlight: espresso
    keep_md: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, fig.keep='all'}
knitr::opts_chunk$set(echo = TRUE)
```

Coursera Reproducible Research Week 2 Course Project
=========================================================================

This submission analyzes the data from a personal activity monitoring device collected during the months of October and November 2012. The data was retrieved from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip. The following describes the measurements and analyses done, and their corresponding outputs. 

```{r showtable, echo=FALSE, results='asis'}
x <- matrix(c(" histogram ", " summary table ", " time series plot ", " summary ", " histogram ", " panel plot " ))

y<- matrix(c(" total number of steps taken each day ", " Mean and median number of steps taken each day ", " average number of steps taken ", " 5-minute interval that, on average, contains the maximum number of steps ", " total number of steps taken each day after missing values are imputed ", " the average number of steps taken per 5-minute interval across weekdays and weekends " ))

   cbind(y,x) -> table1
      as.table(table1) -> table1
         colnames(table1) = c("Measurements/Analyses", "Output")

library(xtable)
   xtable(table1) -> table1

print(table1, type= "html")
```

### Analysis and Results

First, read the csv data, and check some of its properties.

```{r}
read.csv("activity.csv") ->dataset
str(dataset)
```

#### A. Total Number of Steps Taken Each Day
```{r, results='hide'}
as.Date(dataset$date,  "%Y-%m-%d") -> dataset$date
```
``` {r}
suppressPackageStartupMessages(library(ggplot2, quietly = TRUE))
suppressPackageStartupMessages(library(dplyr, quietly = TRUE))

group_by(dataset, date) %>% summarize(steps= sum(steps)) -> totalsum

suppressWarnings(figure1<- (ggplot(totalsum, aes(x=steps)) + geom_histogram(aes(y=..count.., fill=..count..), binwidth = 800)+ theme_bw(base_family="Times") + labs(y= "Frequency", x= "Total Number of Steps Per Day")))

suppressWarnings(print(figure1))
```

#### B. Mean and Median Number of Total Number of Steps Taken Each Day
```{r}
with(dataset, tapply(steps, date, sum)) -> mm
   summary(mm) -> mmb
      print(paste("The mean is", mmb[4], "and the median is", mmb[3], sep=" "), quote = F)
```

#### C. Average Number of Steps Taken Per Day
```{r}
suppressWarnings(figure2 <- (ggplot(dataset, aes(interval, steps)) + geom_line(stat = "summary", fun= "mean", na.rm = TRUE, color="red") + theme_bw(base_family = "Times") + labs(y= "Average Number of Steps Taken Per Day", x= "Interval (5 minutes)")))

suppressWarnings(print(figure2))
```

#### D. 5-minute Interval that, on Average, Contains the Maximum Number of steps
```{r}
with(dataset, tapply(steps, interval, mean, na.rm=T)) -> d
   (max(d) == d) -> true
      names(d[true]) -> f
         paste(f, "minute-interval", "with", d[true], "steps", sep = " ") -> f 

print(noquote(f))
```

#### E. Total Number of Steps Taken Each Day After Missing Values Are Imputed
```{r}
sum(is.na(dataset)) -> e
   paste("Number of NAs is", e, sep = " ") -> e

print(noquote(e))
```
Imputation of Missing Values was done using the code chunk below. Missing values (NA) were replaced by the average of all median values of total steps taken computed by date. This was done, because there are dates wherein all of their values are missing; thus, to more or less preserve uniformity and distribution in totality, the said imputation method was done.
```{r}
suppressPackageStartupMessages(library(dplyr, quietly = TRUE))
   with(dataset, tapply(steps, date, median)) -> med

e2 <- tibble(x=1)

for (i in 1:length(dataset$steps)) {if (is.na(dataset$steps[i]) == TRUE) {add_row(e2,x=    mean(med, na.rm = T))-> e2} else {add_row(e2,x= dataset$steps[i]) -> e2}; e2 -> e3}; e3 -> e3;
e3[2:nrow(e3),] -> e3

   cbind(dataset, e3) -> newdata
      group_by(newdata, date) %>% summarize(x= sum(steps, na.rm = TRUE)) -> totalsum2

suppressWarnings(figure6<- (ggplot(totalsum2, aes(x= x)) + geom_histogram(aes(y=..count.., fill=..count..), binwidth = 800)+ theme_bw(base_family="Times") + labs(y= "Frequency", x= "Total Number of Steps Per Day")))

suppressWarnings(print(figure6))
```
```{r}
with(newdata, tapply(x, date, sum)) -> mm2
   summary(mm2)
```
##### Comparison of Summary of Non-Imputed Dataset and Imputed Dataset
```{r}
summary(mm) -> a; as.vector(a) -> a; summary (mm2) -> b; names(b) -> name; as.vector(b)-> b;
   suppressWarnings(rbind(a, b) -> ab); as.data.frame(ab) -> ab; colnames(ab) = name; rownames(ab) = c("Non-Imputed Data", "Imputed Data");

print(ab)
```

#### F. The Average Number of Steps Taken per 5-minute Interval Across Weekdays and Weekends
```{r}
neww <- tibble(x=0) 
for (i in 1:length(newdata$date)) 
{if (weekdays(newdata$date[i])== "Saturday"){add_row(neww, x= 1) -> neww} 
   else if (weekdays(newdata$date[i]) == "Sunday") {add_row(neww, x= 1)-> neww}
   else {add_row(neww, x= 2)-> neww}; neww -> neww3}; neww3 -> neww3; neww3[2:nrow(neww3),] -> 
   value1; as.data.frame(value1) -> newdata$value

mutate(newdata, week= ifelse(value == 2, "weekday", "weekend")) -> f
   f <- transform(f, week= factor(week))

suppressWarnings(figure7<- (ggplot(f, aes(interval, x)) + geom_line(size=1, stat = "summary", fun="mean", na.rm = TRUE, aes(color=week))+ theme_bw(base_family="Times") + labs(y= "Total Number of Steps", x= ("Interval (5-minute)")) + facet_grid(week~.) + suppressMessages(geom_smooth(linetype= 3, size= 2, method="lm", se=FALSE, col="steelblue"))))

suppressWarnings(print(figure7))
```

